---
title: "Model Fitting for Heat Map Experiment"
format: html
---


```{r appendix-setup, message=FALSE, warning=FALSE, echo = F}
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
library(kableExtra)

load('../../data/stimuli.rda')
load('../../data/data1.rda')
load('../../data/data2.rda')
set.seed(3141)

# Data from Shiny app
library(RSQLite)
conn <- dbConnect(SQLite(), "../../shiny-apps/experiment-heat3d/data/stat218-summer2025.db")
# dbListTables(conn)
blocks <- dbReadTable(conn, "blocks")
exp_results <- dbReadTable(conn, "exp_results")
users <- dbReadTable(conn, "users")
dbDisconnect(conn)

# Solutions
solutions <- readRDS("../../data/solutions.rda")
load("../../data/data1.rda")
load("../../data/data2.rda")

library(tidyverse)
# Pre-processing of results
results <- exp_results %>%

  # Remove all practice trials
  filter(set != "practice") %>%

  # Arrange by user id and trial time
  group_by(user_id) %>%
  arrange(start_time) %>%

  # Identify sequence of trials
  mutate(user_seq = ifelse(user_trial_order > lag(user_trial_order), 0, 1),
         user_seq = ifelse(start_time == min(start_time), 0, user_seq)) %>%
  mutate(user_seq = cumsum(user_seq)) %>%

  # Join with blocks
  left_join(blocks, by = "user_id", relationship = 'many-to-many') %>%

  # Remove blocks that were assigned after the trials started
  group_by(user_id, user_seq) %>%
  filter(system_time < min(start_time)) %>%

  # Get time difference with block and filter for smallest difference
  mutate(time_diff_block = min(start_time) - system_time) %>%
  filter(time_diff_block == min(time_diff_block)) %>%

  # Filter so that only the first completed trial is included
  filter(user_seq == min(user_seq))

# Get user sequences with full completions
full_completions <- results %>%
  group_by(user_id, user_seq, block) %>%
  count() %>%
  filter(n %in% c(16,24))

# Inner join to filter
results <- inner_join(results, full_completions) %>%
  select(-c(time_diff_block, n)) %>%
  ungroup()

# Join with results and filter so that only first completed block is there
results <- left_join(results, solutions) %>%
  ungroup() %>%
  group_by(user_id) %>%
  filter(system_time == min(system_time)) %>%
  ungroup() %>%
  filter(between(as_datetime(system_time), as_date('2025-08-01'), as_date('2025-12-31'))) %>%
  mutate(target_ratio = 100*true_ratio,
         target_size = ifelse(z > 50, 50, z*true_ratio),
         target_diff = z-target_size)

results$pair_id <- factor(results$pair_id)

# All instances of starting the experiment
all_starts <- inner_join(blocks, users) %>%
  filter(!str_detect(tolower(user_unique), 'test,'))



## -------------------------------------------------------------
# Combine users and results, remove all "test" entries
users_clean <- results %>%
  inner_join(users, by = 'user_id', relationship = 'many-to-many') %>%
  select(user_id, user_age:user_unique) %>%
  distinct() %>%
  dplyr::filter(!str_detect(tolower(user_unique), 'test,'))


## -------------------------------------------------------------
# Get in-person users (have at least 1 3dp trial)
users_in_person <- users_clean %>%
  inner_join(results) %>%
  group_by(user_id, media) %>%
  summarise(n = n()) %>%
  filter(media == '3dp' & n > 0) %>%
  select(user_id)


## -------------------------------------------------------------
res_q1 <- results %>%
  inner_join(users_clean, by = 'user_id') %>%
  mutate(q1 = case_when(
    user_larger == 'Both values are the same' ~ 'Equal',
    (user_larger != true_larger) & (user_larger != 'Both values are the same') ~ 'Smaller',
    user_larger != 'Both values are the same' & user_larger == true_larger ~ 'Larger'
  ), correct_label = ifelse(user_larger == true_larger, '*', NA)) %>%
  mutate(q1_label = factor(q1, labels = c('Smaller value\n(or incorrect)',
                                          'Equal', 'Larger value'),
                           levels = c('Smaller', 'Equal', 'Larger'), ordered = T),
         prop = round(100*true_ratio,1),
         facet_label = paste0('Stimuli Pair ', pair_id, ' (', prop, '%)'),
         q1 = factor(q1, levels = c('Smaller', 'Equal', 'Larger'), ordered = F))

res_q1_filtered <- res_q1 %>%
  filter(pair_id != 5) %>%
  group_by(user_id) %>%
  summarize(n_trials = n(),
            n_correct = sum(user_larger == true_larger),
            p.value = pbinom(n_correct, size = n_trials, prob = 2/3, lower.tail = F)) %>%
  ungroup() %>%
  filter(p.value <= 0.05)

res_q2 <- results %>%
  mutate(q2_error = user_slider - target_ratio,
         q2_error_cm = log2(abs(user_slider - target_ratio) + 1/8))
res_q2_filtered <- res_q2 %>% inner_join(res_q1_filtered)



## -------------------------------------------------------------
df_time <- res_q2 %>%
  group_by(user_id, block) %>%
  summarise(total_time = (max(end_time)-min(start_time))/60)

median_time_in_person <- df_time %>% filter(user_id %in% users_in_person$user_id) %>%
  pull(total_time) %>%
  median()

median_time_online <- df_time %>% filter(!(user_id %in% users_in_person$user_id)) %>%
  pull(total_time) %>%
  median()


## -------------------------------------------------------------

df_time_grouped <- df_time %>%
  mutate(Section = ifelse(user_id %in% users_in_person$user_id, 'In-person', 'Online')) %>%
  group_by(Section) %>%
  summarise(Mean = mean(total_time),
            SD = sd(total_time),
            Median = median(total_time))
df_time_total <- df_time %>%
  mutate(Section = 'All Participants') %>%
  group_by(Section) %>%
  summarise(Mean = mean(total_time),
            SD = sd(total_time),
            Median = median(total_time))

## -------------------------------------------------------------
res_q1 <- results %>%
  inner_join(users_clean, by = 'user_id') %>%
  mutate(q1 = case_when(
    user_larger == 'Both values are the same' ~ 'Equal',
    (user_larger != true_larger) & (user_larger != 'Both values are the same') ~ 'Smaller',
    user_larger != 'Both values are the same' & user_larger == true_larger ~ 'Larger'
  ), correct_label = ifelse(user_larger == true_larger, '*', NA)) %>%
  mutate(q1_label = factor(q1, labels = c('Smaller value\n(or incorrect)',
                                          'Equal', 'Larger value'),
                           levels = c('Smaller', 'Equal', 'Larger'), ordered = T),
         prop = round(100*true_ratio,1),
         facet_label = paste0('Stimuli Pair ', pair_id, ' (', prop, '%)'),
         q1 = factor(q1, levels = c('Smaller', 'Equal', 'Larger'), ordered = F))

res_q1_filtered <- res_q1 %>%
  filter(pair_id != 5) %>%
  group_by(user_id) %>%
  summarize(n_trials = n(),
            n_correct = sum(user_larger == true_larger),
            p.value = pbinom(n_correct, size = n_trials, prob = 2/3, lower.tail = F)) %>%
  ungroup() %>%
  filter(p.value <= 0.05)




## ----appendix-setup, message=FALSE, warning=FALSE, echo = F----
library(tidyverse)

```

In this section, we include multiple models for the transparency of model selection from the analysis of Chapter 3. Visual inspections of participant responses indicate that several participants clearly did not follow instructions by either using alternate estimation strategies or by carelessly responding [@fig-q1-eda]. While we should take these cases into consideration, there is no singular solution to suggest a "best" model [@jasoIdentificationCarelessResponding2021; @ward2023]. Our purpose for including these alternate models is to examine model sensitivity to results through various levels of participant and response filtering.

For each model, we exclude stimuli pair 5. In theory, participants who answer this question correctly will select that both values are the same and move the slider to 100. From the included participants, as described in @sec-remove-random-categories, virtually all participants who put the slider at 100 answered the first question correctly. However, nearly 60 percent of responses indicated that the values were the same in the first question but did not move the slider to 100. Several factors could have attributed to this, such as slider sensitivity or participants using alternate estimation strategies. Rather than correcting what we think participants should have answered, we removed this stimuli pair and motivate future research into "just noticeable differences" for 3D heat maps.

```{r}
#| fig-height: 3
#| fig-cap: "Counts of response behavior for Stimuli Pair 5. This pair had identical values, which means that true solutions indicates marking that they were the same value and positioning the slider at 100."
#| fig-scap: "Response behavior for identical-value stimuli pair"
#| label: fig-pair5-issues
infilter.labs <- c('Included Participants', 'Excluded Participants')
names(infilter.labs) <- c(TRUE, FALSE)

res_q2 %>% 
  filter(pair_id == 5) %>% 
  mutate(infilter = user_id %in% res_q2_filtered$user_id) %>% 
  group_by(q1_correct = user_larger == true_larger,
           slider100 = user_slider == 100,
           infilter) %>% 
  count() %>% 
  arrange(infilter, q1_correct) %>% 
  ggplot(mapping = aes(x = slider100, y = n, fill = q1_correct)) + 
  geom_col(position = position_dodge()) + 
  geom_text(aes(label = n, y = n+8), position = position_dodge(width = 1),
            size = 3) + 
  labs(x = 'Slider position', y = 'Count',
       fill = 'Correct solution\nto Q1?') +
  facet_wrap(~infilter, labeller = labeller(infilter = infilter.labs)) + 
  scale_fill_manual(labels = c("No", "Yes"),
                    values = c('#b8b8b8', '#1a80bb')) +
  scale_x_discrete(labels = c('Not at 100', 'At 100')) + 
  theme_bw() + 
  theme(aspect.ratio = 1, legend.position = 'bottom')
```

There are multiple levels of participant and response filtering that we considered. For participant-level filtering, we consider the following levels: (1) no participant filtering and (2) remove suspected random guessers. The process of (2) was described in @sec-remove-random-categories. Filtering individual responses is a straightforward process, where we considered (i) include all responses, and (ii) remove responses with incorrect Q1 solutions.

```{r}
# Format: mod_(participant)_(response)

# All participants
mod_all_all <- lmer(q2_error_cm ~ set*media*pair_id + (1|user_id:block/set:media),
     data = filter(res_q2, pair_id != 5))
mod_all_q1 <- lmer(q2_error_cm ~ set*media*pair_id + (1|user_id:block/set:media),
     data = filter(res_q2, pair_id != 5 & user_larger == true_larger))

# Filtered participants
mod_q1_all <- lmer(q2_error_cm ~ set*media*pair_id + (1|user_id:block/set:media),
     data = filter(res_q2_filtered, pair_id != 5))
mod_q1_q1 <- lmer(q2_error_cm ~ set*media*pair_id + (1|user_id:block/set:media),
     data = filter(res_q2_filtered, pair_id != 5 & user_larger == true_larger))


```




Based on @tbl-all-models-anova, the models produce similar p-values for effect sizes. The p-values for media type are consistently significant across all model specifications, where each model shows 2D charts having higher errors than either of the 3D charts. In contrast, the interaction between set and stimuli pair shows more sensitivity to participant filtering. When all responses are included, this interaction is significant regardless of participant filtering. However, the interaction is not significant when removing incorrect Q1 responses for all participants and is only marginally significant for the filtered set of participants. @fig-all-models-media-int, @fig-all-models-set-int, and @fig-all-models-pair-int show the estimated marginal means of the log error for each model.

\newpage

```{r}
#| eval: true
#| label: tbl-all-models-anova
bind_rows(
  car::Anova(mod_all_all, type = 3, test = 'F') %>% 
  data.frame() %>% 
  janitor::clean_names() %>% 
  rownames_to_column('effect'),
  car::Anova(mod_all_q1, type = 3, test = 'F') %>% 
  data.frame() %>% 
  janitor::clean_names() %>% 
  rownames_to_column('effect'),
  car::Anova(mod_q1_all, type = 3, test = 'F') %>% 
  data.frame() %>% 
  janitor::clean_names() %>% 
  rownames_to_column('effect'),
  car::Anova(mod_q1_q1, type = 3, test = 'F') %>% 
  data.frame() %>% 
  janitor::clean_names() %>% 
  rownames_to_column('effect'),
  .id = 'model'
) %>% 
  filter(effect != '(Intercept)') %>% 
  select(model, effect, pr_f) %>% 
  pivot_wider(names_from = effect, values_from = pr_f) %>% 
  mutate(across(where(is.numeric), round, 3)) %>% 
  mutate(model = case_when(
    model=='1' ~ 'All participants, all responses',
    model=='2' ~ 'All participants, Q1 correct',
    model=='3' ~ 'Filtered participants, all responses',
    model=='4' ~ 'Filtered participants, Q1 correct'
  )) %>% 
  column_to_rownames('model') %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column('Term') %>% 
  kable(caption = 'ANOVA Table p-values for model terms', digits = 3, booktabs = T,
        longtable = T)


# car::Anova(mod_all_all, type = 3, test = 'F')
# car::Anova(mod_all_q1, type = 3, test = 'F')
# car::Anova(mod_q1_all, type = 3, test = 'F')
# car::Anova(mod_q1_q1, type = 3, test = 'F')
```


```{r}
em_all_all <- emmeans(mod_all_all, ~media+set+pair_id, pbkrtest.limit = 5800)
em_all_q1 <- emmeans(mod_all_q1, ~media+set+pair_id, pbkrtest.limit = 5800)
em_q1_all <- emmeans(mod_q1_all, ~media+set+pair_id, pbkrtest.limit = 5800)
em_q1_q1 <- emmeans(mod_q1_q1, ~media+set+pair_id, pbkrtest.limit = 5800)
```

```{r}
#| label: fig-all-models-media-int
#| fig-cap: "Estimated Marginal Means for Participant and Response Filtering (Media)."
#| fig-scap: "Marginal means by media across filtering scenarios"
#| fig-height: 6
#| fig-width: 5.5
#| fig-dpi: 600
bind_rows(
  data.frame(em_all_all),
  data.frame(em_all_q1),
  data.frame(em_q1_all),
  data.frame(em_q1_q1),
  .id = 'model'
) %>% 
  mutate(model = case_when(
    model=='1' ~ 'All participants, all responses',
    model=='2' ~ 'All participants, Q1 correct',
    model=='3' ~ 'Filtered participants, all responses',
    model=='4' ~ 'Filtered participants, Q1 correct'
  )) %>% 
  ggplot(mapping = aes(x = media, y = emmean, color = model)) + 
  geom_point(position = position_dodge(width = 2/3), size = 2/3) + 
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 1/4,
                position = position_dodge(width = 2/3)) + 
  facet_wrap(~set+pair_id, labeller = label_both) + 
  guides(color=guide_legend(nrow=2,byrow=TRUE)) + 
  labs(x = 'Media', y = 'Estimated Marginal Mean', color = '') + 
  theme_bw() + 
  theme(aspect.ratio = 1/2, legend.position = 'bottom', #legend.background = element_rect(color = 'black', linewidth = 1/4),
        strip.text = element_text(size = 6, margin = margin(.05,0,.05,0, 'cm')))
```

```{r}
#| label: fig-all-models-set-int
#| fig-cap: "Estimated Marginal Means for Participant and Response Filtering (Set)."
#| fig-scap: "Marginal means by data set across filtering scenarios"
#| fig-height: 6
#| fig-width: 5.5
#| fig-dpi: 600
bind_rows(
  data.frame(em_all_all),
  data.frame(em_all_q1),
  data.frame(em_q1_all),
  data.frame(em_q1_q1),
  .id = 'model'
) %>% 
  mutate(model = case_when(
    model=='1' ~ 'All participants, all responses',
    model=='2' ~ 'All participants, Q1 correct',
    model=='3' ~ 'Filtered participants, all responses',
    model=='4' ~ 'Filtered participants, Q1 correct'
  )) %>% 
  ggplot(mapping = aes(x = set, y = emmean, color = model)) + 
  geom_point(position = position_dodge(width = 2/3), size = 2/3) + 
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 1/4,
                position = position_dodge(width = 2/3)) + 
  facet_wrap(~media+pair_id, labeller = label_both) + 
  guides(color=guide_legend(nrow=2,byrow=TRUE)) + 
  labs(x = 'Set', y = 'Estimated Marginal Mean', color = '') + 
  theme_bw() + 
  theme(aspect.ratio = 1/2, legend.position = 'bottom', #legend.background = element_rect(color = 'black', linewidth = 1/4),
        strip.text = element_text(size = 6, margin = margin(.05,0,.05,0, 'cm')))
```

```{r}
#| label: fig-all-models-pair-int
#| fig-cap: "Estimated Marginal Means for Participant and Response Filtering (Pair)."
#| fig-scap: "Marginal means by stimuli pair across filtering scenarios"
#| fig-height: 6
#| fig-width: 5.5
#| fig-dpi: 600
bind_rows(
  data.frame(em_all_all),
  data.frame(em_all_q1),
  data.frame(em_q1_all),
  data.frame(em_q1_q1),
  .id = 'model'
) %>% 
  mutate(model = case_when(
    model=='1' ~ 'All participants, all responses',
    model=='2' ~ 'All participants, Q1 correct',
    model=='3' ~ 'Filtered participants, all responses',
    model=='4' ~ 'Filtered participants, Q1 correct'
  )) %>% 
  ggplot(mapping = aes(x = pair_id, y = emmean, color = model)) + 
  geom_point(position = position_dodge(width = 1/2), size = 2/3) + 
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 1/4,
                position = position_dodge(width = 1/2)) + 
  facet_wrap(~media+set, labeller = label_both) + 
  guides(color=guide_legend(nrow=2,byrow=TRUE)) + 
  labs(x = 'Stimuli Pair', y = 'Estimated Marginal Mean', color = '') + 
  theme_bw() + 
  theme(aspect.ratio = 1/2, legend.position = 'bottom', #legend.background = element_rect(color = 'black', linewidth = 1/4),
        strip.text = element_text(size = 6, margin = margin(.05,0,.05,0, 'cm')))
```
